---
title: "Visium SLICE_1"
output: html_notebook
---

```{r}
library(Seurat)
library(ggplot2)
library(cowplot)
library(dplyr)
library(devtools)
library(voxhunt)
library(gplots)
library(dendextend)
library(Matrix)
library(clustree)
library(tidyverse)
library(gtools)
```

load dataframe
```{r}
SLICE_1 <- Load10X_Spatial(data.dir = "/links/groups/treutlein/DATA/sequencing/20200225_P1168_ASHLEY_10X_VISIUM_Lineage_Tracing_Visium_d15/processed/SLICE_1/outs/", slice = "SLICE_1", filter.matrix = TRUE)
```

sctransfrom to normalize the data and detect high-variance features
```{r}
SLICE_1 <- SCTransform(SLICE_1, assay = "Spatial", verbose = FALSE)
```

write table with raw data counts
```{r}
write.table(SLICE_1@assays$SCT@counts, file = "~/cell_lineage/Visium/V19S23-053/SLICE_1/data_out/rawdataSLICE1_SCT.txt", quote = FALSE, sep = "\t")
write.table(SLICE_1@assays$Spatial@counts, file = "~/cell_lineage/Visium/V19S23-053/SLICE_1/data_out/rawdataSLICE1_Spatial.txt", quote = FALSE, sep = "\t")

SCT_cibsortx_input <- read.table("~/cell_lineage/Visium/V19S23-053/SLICE_1/data_out/rawdataSLICE1_SCT.txt", row.names = NULL, sep = "\t")
head(SCT_cibsortx_input)
names(SCT_cibsortx_input)[names(SCT_cibsortx_input) == "row.names"] <- "genes"
head(SCT_cibsortx_input)
write.table(SCT_cibsortx_input, file = "~/cell_lineage/Visium/V19S23-053/SLICE_1/data_out/SCT_cibsortx_input_SLICE_1.txt", quote = FALSE, sep = "\t", row.names=F)

Spatial_cibsortx_input <- read.table("~/cell_lineage/Visium/V19S23-053/SLICE_1/data_out/rawdataSLICE1_Spatial.txt", row.names = NULL, sep = "\t")
head(Spatial_cibsortx_input)
names(Spatial_cibsortx_input)[names(Spatial_cibsortx_input) == "row.names"] <- "genes"
head(Spatial_cibsortx_input)
write.table(Spatial_cibsortx_input, file = "~/cell_lineage/Visium/V19S23-053/SLICE_1/data_out/Spatial_cibsortx_input_SLICE_1.txt", quote = FALSE, sep = "\t", row.names=F)
```

interactive  plot to identify spots not connected to tissue
```{r}
ISpatialDimPlot(SLICE_1, image = NULL, alpha = c(0.3, 1))
```

Filter out spots with less than 1000 feature and not connected to tissue
```{r}
hist(SLICE_1$nFeature_Spatial)
low_feature_cut <- 1000
WhichCells(SLICE_1, expression = nFeature_Spatial < low_feature_cut)
SpatialDimPlot(SLICE_1, cells.highlight = WhichCells(SLICE_1, expression = nFeature_Spatial < low_feature_cut))

cells_drop <- WhichCells(SLICE_1, expression = nFeature_Spatial < low_feature_cut)
all_cells_todrop <- rownames(SLICE_1@meta.data)[which(rownames(SLICE_1@meta.data) %in% c("CTTAGCCCGGATAGTG-1",cells_drop))] #"CTTAGCCCGGATAGTG-1" is the only spot with >1000 genes but not touching tissue

png("~/cell_lineage/Visium/V19S23-053/SLICE_1/plot_out/Dimplot_cellstodrop.png", width = 10, height = 10, units = "in", res = 300)
SpatialDimPlot(SLICE_1, cells.highlight = all_cells_todrop, pt.size.factor = 2, stroke = .5)
dev.off()

cells_to_keep <- rownames(SLICE_1@meta.data)[!rownames(SLICE_1@meta.data) %in% all_cells_todrop]

SLICE_1_sub <- subset(SLICE_1, cells = cells_to_keep)
```

```{r}
SLICE_1_sub <- FindVariableFeatures(SLICE_1_sub, selection.method = "vst", nfeatures = 2000)
```

Dimensionality reduction, clustering, and visualization
```{r}
SLICE_1_sub <- RunPCA(SLICE_1_sub, assay = "SCT", features = VariableFeatures(object = SLICE_1_sub), verbose = FALSE)
ElbowPlot(SLICE_1_sub)
```

Find 
```{r}
dims = 1:8
SLICE_1_sub <- FindNeighbors(SLICE_1_sub, reduction = "pca", dims = dims)
```

Find best resolution for clustering
```{r}
SLICE_1_sub <- FindClusters(object = SLICE_1_sub, verbose = T, resolution = 0.1)
SLICE_1_sub <- FindClusters(object = SLICE_1_sub, verbose = T, resolution = 0.3)
SLICE_1_sub <- FindClusters(object = SLICE_1_sub, verbose = T, resolution = 0.5)
SLICE_1_sub <- FindClusters(object = SLICE_1_sub, verbose = T, resolution = 0.7)

# Make plot 
clus.tree.out <- clustree(SLICE_1_sub) +
    theme(legend.position = "top") + 
    scale_color_brewer(palette = "Set1") +
    scale_edge_color_continuous(low = "grey80", high = "red")
# Plot 
png("~/cell_lineage/Visium/V19S23-053/SLICE_1/plot_out/resolution_clustree.png", width = 15, height = 10, units = "in", res = 300)
clus.tree.out
dev.off()
```

Find clusters and run UMAP
```{r}
res <- 0.5
SLICE_1_sub <- FindClusters(SLICE_1_sub, verbose = FALSE, resolution = res)
SLICE_1_sub <- RunUMAP(SLICE_1_sub, reduction = "pca", dims = dims)
```

plot clusters
```{r}
p1 <- DimPlot(SLICE_1_sub, reduction = "umap", label = TRUE, pt.size = 2, )
p2 <- SpatialDimPlot(SLICE_1_sub, label = TRUE, label.size = 3, pt.size.factor = 2.5, stroke = 0.5)
p1_2 <- plot_grid(p1, p2)

p1_2
ggsave(p1_2, filename = "~/cell_lineage/Visium/V19S23-053/SLICE_1/plot_out/clusters.png", height = 10, width = 20, units = "in", dpi = 300)
```

plot nfeatures and ncounts
```{r}
plot3 <- VlnPlot(SLICE_1_sub, features = "nCount_SCT", pt.size = 0.1) + NoLegend()
plot4 <- SpatialFeaturePlot(SLICE_1_sub, features = "nCount_SCT", pt.size.factor = 2.5, stroke = 0.5) + theme(legend.position = "right")
p3_p4 <- plot_grid(plot3, plot4)


plot5 <- VlnPlot(SLICE_1_sub, features = "nFeature_SCT", pt.size = 0.1) + NoLegend()
plot6 <- SpatialFeaturePlot(SLICE_1_sub, features = "nFeature_SCT", pt.size.factor = 2.5, stroke = 0.5) + theme(legend.position = "right")
p5_p6 <- plot_grid(plot5, plot6)

p3_p4
p5_p6

ggsave(p3_p4, filename = "~/cell_lineage/Visium/V19S23-053/SLICE_1/plot_out/nCountSCT.png", height = 10, width = 20, units = "in", dpi = 300)

ggsave(p5_p6, filename = "~/cell_lineage/Visium/V19S23-053/SLICE_1/plot_out/nFeatureSCT.png", height = 10, width = 20, units = "in", dpi = 300)
```

find DE genes and plot top 2
```{r}
de_markers <- FindAllMarkers(SLICE_1_sub)
de_markers %>% group_by(cluster) %>% top_n(20, avg_logFC)

top_2_ea_clus <- de_markers %>% group_by(cluster) %>% top_n(2, avg_logFC)

png(filename = "~/cell_lineage/Visium/V19S23-053/SLICE_1/plot_out/top2de.png", width = 20, height = 15, units = "in", res = 300)
SpatialFeaturePlot(object = SLICE_1_sub, features = top_2_ea_clus$gene, ncol = 5, pt.size.factor = 2.5, stroke = 0.5)
dev.off()

write.csv(de_markers, file = "~/cell_lineage/Visium/V19S23-053/SLICE_1/data_out/de_markers_bycluster.csv")
de_markers <- read.csv(file = "~/cell_lineage/Visium/V19S23-053/SLICE_1/data_out/de_markers_bycluster.csv")


```

Voxhunt
```{r}
theme_set(theme_bw())
load_aba_data('/links/groups/treutlein/PUBLIC_DATA/tools/voxhunt_package/')
head(Idents(SLICE_1_sub), 5)
head(SLICE_1_sub@meta.data)

regional_markers <- structure_markers('E13') %>%
    group_by(group) %>%
    top_n(10, auc) %>% 
    {unique(.$gene)}
regional_markers
vox_map <- voxel_map(SLICE_1_sub, stage = 'E13', group_name = 'seurat_clusters', genes_use = regional_markers)
vox_map

png(filename = "~/cell_lineage/Visium/V19S23-053/SLICE_1/plot_out/voxhunt.png", height = 7, width = 7, units = "in", res = 300)
plot_map(vox_map)
dev.off()
```

Feature Plots
```{r}
png(filename = "~/cell_lineage/Visium/V19S23-053/SLICE_1/plot_out/markergenes.png", width = 20, height = 15, units = "in", res = 300)
SpatialFeaturePlot(SLICE_1_sub, features = c("Tomato","OTX2", 'PAX6', "TBR1", "NEUROD6", "FOXG1", "HOXB2", "HOXB8", "HOXB5", "DCN", "TTR", "PRPH", "NEUROD1", "SOX10", "PMEL", "RSPO2", "RSPO3", "MEIS1"), ncol = 6, pt.size.factor = 2.5, stroke = 0.5)
dev.off()
```

Calculate NPC and Neuron Scores
```{r}
NPC_tab <- read.table(file = "~/cell_lineage/plot_out/SLICE1_plots/list.NPC_signature.txt")
Neuron_tab <- read.table(file = "~/cell_lineage/plot_out/SLICE1_plots/list.neuron_signature.txt")

NPC_sub <- as.data.frame(colSums(SLICE_1_sub@assays$SCT@data[NPC_tab$V1, ]))
Neuron_sub <- as.data.frame(colSums(SLICE_1_sub@assays$SCT@data[Neuron_tab$V1, ]))
NPC_Neuron <- cbind(NPC_sub, Neuron_sub)
colnames(NPC_Neuron) <- c("NPC_Score", "Neuron_Score")

SLICE_1_sub@meta.data <- cbind(SLICE_1_sub@meta.data, NPC_Neuron)

png(filename = "~/cell_lineage/Visium/V19S23-053/SLICE_1/plot_out/npc_neuronal_scores.png", width = 10, height = 5, units = "in", res = 300)
SpatialFeaturePlot(SLICE_1_sub, features = c("NPC_Score","Neuron_Score"), pt.size.factor = 2.5, stroke = 0.5)
dev.off()
```

ashley try intersect with DE from clusters
```{r}
# SLICE_1_sub <- FindSpatiallyVariableFeatures(SLICE_1_sub, assay = "SCT", features = VariableFeatures(SLICE_1_sub)[1:2000], selection.method = "markvariogram")
# 
# head(SpatiallyVariableFeatures(SLICE_1_sub, selection.method = "markvariogram"), 50)
# top.features <- head(SpatiallyVariableFeatures(SLICE_1_sub, selection.method = "markvariogram"), )
# clus
# 
# SpatialFeaturePlot(SLICE_1_sub, features = top.features, ncol = 3, alpha = c(0.1, 1))
```

create table for 3D mapping
```{r}
# SLICE1_metadata <- SLICE_1_sub@images$SLICE_1_sub@coordinates
# SLICE1_metadata$cell_id <- rownames(SLICE1_metadata)
# 
# SLICE_1_sub@meta.data$cell_id <- rownames(SLICE_1_sub@meta.data)
# SLICE1_metadata1 <- left_join(SLICE1_metadata, SLICE_1_sub@meta.data)
# rownames(SLICE1_metadata1) <- SLICE1_metadata1$cell_id
# 
# markers <- c("TBR1", "NEUROD6", "FOXG1","Tomato","OTX2", 'PAX6',"HOXB2", "HOXB8", "HOXB5", "DCN", "TTR", "PRPH", "NEUROD1", "SOX10", "PMEL", "RSPO2", "RSPO3", "MEIS1")
# 
# data_expsubSLICE1 <- SLICE_1_sub@assays$SCT@data[markers, ]
# 
# write.csv(x = data_expsubSLICE1, file = "~/cell_lineage/plot_out/SLICE1_plots/SCT_data_expsubSLICE1.csv")
```

```{r}
save(SLICE_1, file = "~/cell_lineage/Visium/V19S23-053/SLICE_1/SLICE_1.RData")
save(SLICE_1_sub, file = "~/cell_lineage/Visium/V19S23-053/SLICE_1/SLICE_1_sub.RData")
```

```{r}
SLICE_1_sub_coords <- SLICE_1_sub@images$SLICE_1@coordinates
SLICE_1_sub_anno <- cbind(SLICE_1_sub@meta.data, SLICE_1_sub_coords)
write.csv(SLICE_1_sub_anno, file = "~/cell_lineage/Visium/V19S23-053/SLICE_1/data_out/SLICE_1_anno_w_coords.csv")
```



