---
title: "SLICE_3"
output: html_notebook
---

```{r}
library(Seurat)
library(ggplot2)
library(cowplot)
library(dplyr)
library(devtools)
```

```{r}
SLICE_3 <- Load10X_Spatial(data.dir = "/local1/DATA/sequencing/20200225_P1168_ASHLEY_10X_VISIUM_Lineage_Tracing_Visium_d15/processed/manual_alignments/SLICE_3/outs/", slice = "SLICE_3", filter.matrix = TRUE)
```

sctransfrom to normalize the data and detect high-variance features
```{r}
SLICE_3 <- SCTransform(SLICE_3, assay = "Spatial", verbose = FALSE)
```

write table with raw data counts
```{r}
write.table(SLICE_3@assays$SCT@counts, file = "~/cell_lineage/Visium/V19S23-053/SLICE_3/data_out/rawdataSLICE3_SCT.txt", quote = FALSE, sep = "\t")
write.table(SLICE_3@assays$Spatial@counts, file = "~/cell_lineage/Visium/V19S23-053/SLICE_3/data_out/rawdataSLICE3_Spatial.txt", quote = FALSE, sep = "\t")

SCT_cibsortx_input <- read.table("~/cell_lineage/Visium/V19S23-053/SLICE_3/data_out/rawdataSLICE3_SCT.txt", row.names = NULL, sep = "\t")
head(SCT_cibsortx_input)
names(SCT_cibsortx_input)[names(SCT_cibsortx_input) == "row.names"] <- "genes"
head(SCT_cibsortx_input)
write.table(SCT_cibsortx_input, file = "~/cell_lineage/Visium/V19S23-053/SLICE_3/data_out/SCT_cibsortx_input_SLICE_3.txt", quote = FALSE, sep = "\t", row.names=F)

Spatial_cibsortx_input <- read.table("~/cell_lineage/Visium/V19S23-053/SLICE_3/data_out/rawdataSLICE3_Spatial.txt", row.names = NULL, sep = "\t")
head(Spatial_cibsortx_input)
names(Spatial_cibsortx_input)[names(Spatial_cibsortx_input) == "row.names"] <- "genes"
head(Spatial_cibsortx_input)
write.table(Spatial_cibsortx_input, file = "~/cell_lineage/Visium/V19S23-053/SLICE_3/data_out/Spatial_cibsortx_input_SLICE_3.txt", quote = FALSE, sep = "\t", row.names=F)
```

interactive  plot to identify spots not connected to tissue
```{r}
ISpatialDimPlot(SLICE_3, image = NULL, alpha = c(0.3, 1))
```

Filter out spots with less than 1000 feature and not connected to tissue
```{r}
# hist(SLICE_3$nFeature_Spatial)
# low_feature_cut <- 1000
# WhichCells(SLICE_3, expression = nFeature_Spatial < low_feature_cut)
# SpatialDimPlot(SLICE_3, cells.highlight = WhichCells(SLICE_3, expression = nFeature_Spatial < low_feature_cut))
# SpatialFeaturePlot(SLICE_3, features = 'nFeature_SCT')

# cells_drop <- WhichCells(SLICE_3, expression = nFeature_Spatial < low_feature_cut)
# all_cells_todrop <- rownames(SLICE_3@meta.data)[which(rownames(SLICE_3@meta.data) %in% c(cells_drop))] #all spots are touching tissue (this slice is manually tissue annotated upstream in loupe browser)
# 
# png("~/cell_lineage/Visium/V19S23-053/SLICE_3/plot_out/Dimplot_cellstodrop.png", width = 10, height = 10, units = "in", res = 300)
# SpatialDimPlot(SLICE_3, cells.highlight = all_cells_todrop, pt.size.factor = 2, stroke = .5)
# dev.off()
# 
# cells_to_keep <- rownames(SLICE_3@meta.data)[!rownames(SLICE_3@meta.data) %in% all_cells_todrop]
# 
# SLICE_3_sub <- subset(SLICE_3, cells = cells_to_keep)
```

```{r}
SLICE_3 <- FindVariableFeatures(SLICE_3, selection.method = "vst", nfeatures = 2000)
```

Dimensionality reduction, clustering, and visualization
```{r}
SLICE_3 <- RunPCA(SLICE_3, assay = "SCT", features = VariableFeatures(object = SLICE_3), verbose = FALSE)
ElbowPlot(SLICE_3)
```

Find 
```{r}
dims = 1:9
SLICE_3 <- FindNeighbors(SLICE_3, reduction = "pca", dims = dims)
```

Find best resolution for clustering
```{r}
SLICE_3 <- FindClusters(object = SLICE_3, verbose = T, resolution = 0.1)
SLICE_3 <- FindClusters(object = SLICE_3, verbose = T, resolution = 0.3)
SLICE_3 <- FindClusters(object = SLICE_3, verbose = T, resolution = 0.5)
SLICE_3 <- FindClusters(object = SLICE_3, verbose = T, resolution = 0.7)

# Make plot 
clus.tree.out <- clustree(SLICE_3) +
    theme(legend.position = "top") + 
    scale_color_brewer(palette = "Set1") +
    scale_edge_color_continuous(low = "grey80", high = "red")
# Plot 
png("~/cell_lineage/Visium/V19S23-053/SLICE_3/plot_out/resolution_clustree.png", width = 15, height = 10, units = "in", res = 300)
clus.tree.out
dev.off()
```

Find clusters and run UMAP
```{r}
res <- 0.5
SLICE_3 <- FindClusters(SLICE_3, verbose = FALSE, resolution = res)
SLICE_3 <- RunUMAP(SLICE_3, reduction = "pca", dims = dims)
```

plot clusters
```{r}
p1 <- DimPlot(SLICE_3, reduction = "umap", label = TRUE, pt.size = 2, )
p2 <- SpatialDimPlot(SLICE_3, label = TRUE, label.size = 3, pt.size.factor = 2.5, stroke = 0.5)
p1_2 <- plot_grid(p1, p2)

p1_2
ggsave(p1_2, filename = "~/cell_lineage/Visium/V19S23-053/SLICE_3/plot_out/clusters.png", height = 10, width = 20, units = "in", dpi = 300)
```

plot nfeatures and ncounts
```{r}
plot3 <- VlnPlot(SLICE_3, features = "nCount_SCT", pt.size = 0.1) + NoLegend()
plot4 <- SpatialFeaturePlot(SLICE_3, features = "nCount_SCT", pt.size.factor = 2.5, stroke = 0.5) + theme(legend.position = "right")
p3_p4 <- plot_grid(plot3, plot4)


plot5 <- VlnPlot(SLICE_3, features = "nFeature_SCT", pt.size = 0.1) + NoLegend()
plot6 <- SpatialFeaturePlot(SLICE_3, features = "nFeature_SCT", pt.size.factor = 2.5, stroke = 0.5) + theme(legend.position = "right")
p5_p6 <- plot_grid(plot5, plot6)

p3_p4
p5_p6

ggsave(p3_p4, filename = "~/cell_lineage/Visium/V19S23-053/SLICE_3/plot_out/nCountSCT.png", height = 10, width = 20, units = "in", dpi = 300)

ggsave(p5_p6, filename = "~/cell_lineage/Visium/V19S23-053/SLICE_3/plot_out/nFeatureSCT.png", height = 10, width = 20, units = "in", dpi = 300)
```

```{r}
plot7 <- VlnPlot(SLICE_3, features = "nCount_SCT", pt.size = 0.1) + NoLegend()
plot8 <- SpatialFeaturePlot(SLICE_3, features = "nCount_SCT", pt.size.factor = 2.5) + theme(legend.position = "right")
p7_p8 <- plot_grid(plot7, plot8)


plot9 <- VlnPlot(SLICE_3, features = "nFeature_Spatial", pt.size = 0.1) + NoLegend()
plot10 <- SpatialFeaturePlot(SLICE_3, features = "nFeature_Spatial", pt.size.factor = 2.5) + theme(legend.position = "right")
p9_p10 <- plot_grid(plot9, plot10)

p7_p8
p9_p10

ggsave(p7_p8, filename = "~/cell_lineage/Visium/V19S23-053/SLICE_3/plot_out/clusters_nCountSCT.pdf", height = 7, width = 16, units = "in")
ggsave(p9_p10, filename = "~/cell_lineage/Visium/V19S23-053/SLICE_3/plot_out/clusters_nFeaturesSCT.pdf", height = 7, width = 16, units = "in")
```

Highlight cells
```{r}
SpatialDimPlot(SLICE_3, cells.highlight = CellsByIdentities(object = SLICE_3, idents = c(0, 1, 2, 3, 4, 5, 6)), facet.highlight = TRUE, ncol = 3, pt.size.factor = 2.5)
```

Calculate NPC and Neuron Scores
```{r}
NPC_tab <- read.table(file = "~/cell_lineage/plot_out/SLICE1_plots/list.NPC_signature.txt")
Neuron_tab <- read.table(file = "~/cell_lineage/plot_out/SLICE1_plots/list.neuron_signature.txt")

NPC_sub <- as.data.frame(colSums(SLICE_3@assays$SCT@data[NPC_tab$V1, ]))
Neuron_sub <- as.data.frame(colSums(SLICE_3@assays$SCT@data[Neuron_tab$V1, ]))
NPC_Neuron <- cbind(NPC_sub, Neuron_sub)
colnames(NPC_Neuron) <- c("NPC_Score", "Neuron_Score")

SLICE_3@meta.data <- cbind(SLICE_3@meta.data, NPC_Neuron)

png(filename = "~/cell_lineage/Visium/V19S23-053/SLICE_3/plot_out/npc_neuron_scores.png", width = 10, height = 7, units = "in", res = 300)
SpatialFeaturePlot(SLICE_3, features = c("NPC_Score","Neuron_Score"), pt.size.factor = 3, stroke = 1)
dev.off()
```


```{r}
de_markers <- FindAllMarkers(SLICE_3)
de_markers %>% group_by(cluster) %>% top_n(20, avg_logFC)
write.csv(x = de_markers, file = "~/cell_lineage/Visium/V19S23-053/SLICE_3/data_out/de_markers_SLICE_3.csv")

SpatialFeaturePlot(object = SLICE_3, features = rownames(de_markers)[1:3], alpha = c(0.1, 1), ncol = 3, pt.size.factor = 2.5)
```

Voxhunt
```{r}
theme_set(theme_bw())
load_aba_data('/links/groups/treutlein/PUBLIC_DATA/tools/voxhunt_package/')
head(Idents(SLICE_3), 5)
head(SLICE_3@meta.data)

regional_markers <- structure_markers('E13') %>%
    group_by(group) %>%
    top_n(10, auc) %>% 
    {unique(.$gene)}
regional_markers

vox_map <- voxel_map(SLICE_3, stage = 'E13', group_name = 'seurat_clusters', genes_use = regional_markers)
vox_map
pdf(file= "~/cell_lineage/Visium/V19S23-053/SLICE_3/plot_out/voxhunt.pdf")
p <- plot_map(vox_map)
p
dev.off()
p
```

Feature Plots
```{r}
pdf(file = "~/cell_lineage/Visium/V19S23-053/SLICE_3/plot_out/featureplots.pdf", width = 16, height = 7)
SpatialFeaturePlot(SLICE_3, features = c("Tomato","OTX2", 'PAX6'), ncol = 3, pt.size.factor = 2.5)
SpatialFeaturePlot(SLICE_3, features = c("TBR1", "NEUROD6", "FOXG1"), ncol = 3, pt.size.factor = 2.5)
SpatialFeaturePlot(SLICE_3, features = c("HOXB2", "HOXB8", "HOXB5"), ncol = 3, pt.size.factor = 2.5)
SpatialFeaturePlot(SLICE_3, features = c("DCN", "TTR", "PRPH"), ncol = 3, pt.size.factor = 2.5)
SpatialFeaturePlot(SLICE_3, features = c("NEUROD1", "SOX10", "PMEL"), ncol = 3, pt.size.factor = 2.5)
SpatialFeaturePlot(SLICE_3, features = c("RSPO2", "RSPO3", "MEIS1"), ncol = 3, pt.size.factor = 2.5)
dev.off()
```

```{r}
save(SLICE_3, file = "~/cell_lineage/Visium/V19S23-053/SLICE_3/SLICE_3.RData")
```

```{r}
SLICE_3_coords <- SLICE_3@images$SLICE_3@coordinates
SLICE_3_anno <- cbind(SLICE_3@meta.data, SLICE_3_coords)
write.csv(SLICE_3_anno, file = "~/cell_lineage/Visium/V19S23-053/SLICE_3/data_out/SLICE_3_anno_w_coords.csv")
```
