---
title: "Barcode Plotting"
output: html_notebook
---

load data
```{r}
load(file = "~/cell_lineage/Visium/V19S23-053/SLICE_2/SLICE_2_dienmidcibx_.RData")
load(file = "~/cell_lineage/Visium/V19S23-053/SLICE_1/SLICE_1_dienmidcibx_.RData")
load(file = "~/cell_lineage/Visium/V19S23-053/SLICE_3/SLICE_3_dienmidcibx_.RData")
visium_barcodes <- readRDS(file = "~/cell_lineage/Visium/V19S23-053/VISIUM_3D_meta_new.RDS")
write.csv(visium_barcodes, "~/cell_lineage/Visium/V19S23-053/VISIUM_3D_meta_new1.csv")
rownames(visium_barcodes) <- visium_barcodes$X
```

binarize the barcodes, find total barcodes per spot and identify all spots with a barcode
```{r}
visium_barcodes_bi <- as.data.frame(ifelse(test = visium_barcodes[,40:71]=="negative", yes = 0, no = 1))
visium_barcodes_bi$barcodesum <- rowSums(visium_barcodes_bi)
visium_barcodes_bi$hasbarcode <- visium_barcodes_bi$barcodesum
visium_barcodes_bi$hasbarcode <- ifelse(test = visium_barcodes_bi$barcodesum > 0, yes = "yes", no = "no")

rownames(visium_barcodes_bi) <- rownames(visium_barcodes)
visium_barcodes_bi$X <- rownames(visium_barcodes_bi)

table(visium_barcodes_bi$barcodesum)
table(visium_barcodes_bi$hasbarcode) #850 cells have a barcode 

visium_barcodes_bi <- left_join(x = visium_barcodes_bi, y = visium_barcodes, by = "X")

#plot
#number of barcodes by Z
visium_barcodes_bi$Z <- factor(visium_barcodes_bi$Z, levels = c(3, 2, 1))
pdf("~/cell_lineage/Visium/V19S23-053/spots_with_barcodes1.pdf")
ggplot(visium_barcodes_bi, aes(x = Z, group = hasbarcode)) + geom_bar(aes(fill=hasbarcode), position = "fill")
ggplot(visium_barcodes_bi, aes(x = Z, group = hasbarcode)) + geom_bar(aes(fill=hasbarcode)) + theme_classic()
table(visium_barcodes_bi$hasbarcode, visium_barcodes_bi$Z)

ggplot(visium_barcodes_bi, aes(x = spotassignment, group = hasbarcode)) + geom_bar(aes(fill=hasbarcode), position = "fill")

visium_barcodes_bi$spotassignment <- factor(visium_barcodes_bi$spotassignment, levels = c("dien.mid","low_maxscore","hind","mesen","pns","tel", "epi","schwann"))
ggplot(visium_barcodes_bi, aes(x = spotassignment, group = hasbarcode)) + geom_bar(aes(fill=hasbarcode))
dev.off()

table(visium_barcodes_bi$hasbarcode, visium_barcodes_bi$Z)
table(visium_barcodes_bi$hasbarcode, visium_barcodes_bi$spotassignment)

pdf("~/cell_lineage/Visium/V19S23-053/barcodes_presencevabsence3D.pdf")
scatter3D(x = visium_barcodes_bi$row.sc, y = visium_barcodes_bi$col.sc,
          z = as.numeric(visium_barcodes_bi$ZZ),
          colvar = as.numeric(as.factor(visium_barcodes_bi$hasbarcode)),
          col = c("#DADCDC", "#8B0000"),
          main = "Barcodes", bty = "n", pch = 19, cex = 1.0, theta = 75, phi = 30, colkey = FALSE)
dev.off()
```

Plot individual example barcodes
```{r}
pdf("~/cell_lineage/Visium/V19S23-053/barcodes3D.pdf")
scatter3D(x = visium_barcodes_bi$row.sc, y = visium_barcodes_bi$col.sc,
          z = as.numeric(visium_barcodes_bi$ZZ),
          colvar = as.numeric(as.factor(visium_barcodes_bi$ACGCCGCGCTA.y)),
          col = c("#0893CE", "#DADCDC"),
          main = "ACGCCGCGCTA", bty = "n", pch = 19, cex = 1.0, theta = 75, phi = 30, colkey = FALSE)

scatter3D(x = visium_barcodes_bi$row.sc, y = visium_barcodes_bi$col.sc,
          z = as.numeric(visium_barcodes_bi$ZZ),
          colvar = as.numeric(as.factor(visium_barcodes_bi$TGGTTGGGAGA.y)),
          col = c("#DADCDC", "#0893CE"),
          main = "TGGTTGGGAGA", bty = "n", pch = 19, cex = 1.0, theta = 75, phi = 30, colkey = FALSE)

scatter3D(x = visium_barcodes_bi$row.sc, y = visium_barcodes_bi$col.sc,
          z = as.numeric(visium_barcodes_bi$ZZ),
          colvar = as.numeric(as.factor(visium_barcodes_bi$AAAGGGGAGTA.y)),
          col = c("#0893CE", "#DADCDC"),
          main = "AAAGGGGAGTA", bty = "n", pch = 19, cex = 1.0, theta = 75, phi = 30, colkey = FALSE)

scatter3D(x = visium_barcodes_bi$row.sc, y = visium_barcodes_bi$col.sc,
          z = as.numeric(visium_barcodes_bi$ZZ),
          colvar = as.numeric(as.factor(visium_barcodes_bi$TCAAAGGTATA.y)),
          col = c("#DADCDC", "#0893CE"),
          main = "TCAAAGGTATA", bty = "n", pch = 19, cex = 1.0, theta = 75, phi = 30, colkey = FALSE)

scatter3D(x = visium_barcodes_bi$row.sc, y = visium_barcodes_bi$col.sc,
          z = as.numeric(visium_barcodes_bi$ZZ),
          colvar = as.numeric(as.factor(visium_barcodes_bi$TTGTGTCTAGT.y)),
          col = c("#DADCDC", "#0893CE"),
          main = "TTGTGTCTAGT", bty = "n", pch = 19, cex = 1.0, theta = 75, phi = 30, colkey = FALSE)
dev.off()
```

read in file where we combine barcodes for overlap (specific barcodes are detailed below, you can combine any barcodes you like we choose 4 barcodes) (manualy created from visium_barcode_bi file)
```{r}
Visum.meta.all.barcodes1 <- read.csv("~/cell_lineage/Visium/V19S23-053/visum_metaallbarcodes1.csv") #after combining example barcodes 
```

```{r}
levels(as.factor(Visum.meta.all.barcodes1$example_barcodes3))
#F7F747 -yellow - AAAGGGGAGTA #1
#0893CE - blue - ACGCCGCGCTA # 2
#DADCDC - grey - negative #4
#D50606 - red - TCAAAGGTATA #5
#E48F21 - orange - TCAAAGGTATA/AAAGGGGAGTA #6
#13A226 - green -ACGCCGCGCTA/AAAGGGGAGTA #3
#902ABC - purple- TTTGTGAGGAA

pdf("~/cell_lineage/Visium/V19S23-053/barcodes_combo4bars.pdf")
scatter3D(x = Visum.meta.all.barcodes1$row.sc, y = Visum.meta.all.barcodes1$col.sc,
          z = as.numeric(Visum.meta.all.barcodes1$ZZ),
          colvar = as.numeric(as.factor(Visum.meta.all.barcodes1$example_barcodes3)),
          col = c("#F7F747", "#0893CE", "#13A226", "#DADCDC", "#D50606", "#E48F21", "#902ABC"),
          bty = "n", pch = 19, cex = 1.0, theta = 75, phi = 30, colkey = FALSE)
dev.off()
``` 

Plot Total Barcodes per Spot (3D)
```{r}
pdf("~/cell_lineage/Visium/V19S23-053/barcodes_total.pdf")
scatter3D(x = visium_barcodes_bi$row.sc, y = visium_barcodes_bi$col.sc,
          z = as.numeric(visium_barcodes_bi$ZZ),
          colvar = as.numeric(as.factor(visium_barcodes_bi$barcodesum)),
          col = ramp.col(c("white", "red")),
          bty = "n", pch = 19, cex = 1.0, theta = 75, phi = 30, colkey = TRUE)
dev.off()

table(visium_barcodes_bi$barcodesum)
```

```{r}
write.csv(visium_barcodes_bi, file = "~/cell_lineage/Visium/V19S23-053/visium_barcodesbi.csv")
```

find % mito per spot
```{r}
SLICE_1_sub
SLICE_2_sub
SLICE_3
x <- PercentageFeatureSet(SLICE_1_sub, pattern = "^MT-")
y <- PercentageFeatureSet(SLICE_2_sub, pattern = "^MT-")
z <- PercentageFeatureSet(SLICE_3, pattern = "^MT-")
rownames(x) <- gsub(pattern = "-1", replacement = "_S1", x = rownames(x))
rownames(y) <- gsub(pattern = "-1", replacement = "_S2", x = rownames(y))
rownames(z) <- gsub(pattern = "-1", replacement = "_S3", x = rownames(z))
w <- rbind(x, y)
v <- rbind (w, z)
mean(v$nCount_SCT)
v$spot_id <- rownames(v)
colnames(v) <- c("percent.mito", "spot_id")
```

construct final metadata
```{r}
colnames(visium_barcodes_bi)
colnames(visium_barcodes)
colnames(Visum.meta.all.barcodes1)

metadata_visium <- visium_barcodes_bi[,c(33:35, 37:40, 48:59, 62, 64:72)]
metadata_visium <- left_join(metadata_visium, visium_barcodes[, c(1, 40:71)])
metadata_visium <- left_join(metadata_visium, Visum.meta.all.barcodes1[,c(1,36)])
colnames(metadata_visium)
names(metadata_visium)[3] <- "spot_id"
rownames(metadata_visium) <- metadata_visium$spot_id
metadata_visium <- left_join(metadata_visium, v)

write.csv(metadata_visium, file = "~/cell_lineage/Visium/V19S23-053/metadata_visium.csv")
```




